---
title: "Instruction and an Example of BSS Algorithm"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setting Up

Before implementing the algorithm, we need the following set up:

* Install required packages and load packages.
* Call functions from .R file and .cpp files
    + "Functions_BSS.R"
    + "Functions_BSS.cpp"

```{r,eval=FALSE}
install.packages("mvtnorm")
install.packages("lattice")
install.packages("igraph")
install.packages("pracma")
install.packages("Rcpp")
install.packages("RcppArmadillo")
```

```{r}
library("mvtnorm")
library("lattice")
library("igraph")
library("pracma")
library("Rcpp")
library("RcppArmadillo")
```

```{r}
source("Functions_BSS.R")
#cpp code for block fused lasso and block lasso
sourceCpp("Functions_BSS.cpp")
```


## Simulation Scenario Setting

In order to generating the VAR time series data, the following arguments are required:

* $\texttt{T}$: the number of time points
* $\texttt{p}$: the number of time series components
* $\texttt{brk}$: the true break points with $T+1$ as the last element
* $\texttt{m0}$: the true number of break points
* $\texttt{q.t}$: the true AR order
* $\texttt{phi.full}$: the true AR coefficient matrix

Here is an example setting of Scenario A.2 in the manuscript. In this setting, $T = 4,000$, $p = 10$, $q.t = 1$, $m0 = 2$, two break points $t_1 = \lfloor \frac{T}{3} \rfloor = 1333$, $t_2 =  \lfloor \frac{2T}{3} \rfloor  = 2666$ and the autoregressive coefficients are chosen to have the random structure and entries.


```{r}
########## General Settings ######################
T <- (10^3*4); #number of time points
p <- 10; # number of time series components
brk <- c(floor(T/3),floor(2*T/3),T+1); # true break points with T+1 as the last element
m0 <- length(brk) -1; # number of break points
q.t <- 1; # the true AR order
m <- m0+1 #number of segments
sp_density <- c(0.05, 0.05, 0.05) #sparsity level (5%)
```

## Data Generation

After setting the general arguments described above, we can now generate the time series data by using the function  $\texttt{simu_var()}$ as follows. Here, the covariance matrix of the noise process $\Sigma_\varepsilon = I_T$ (where $I_T$ stands for the identity matrix). 

```{r}
#randomly generate transition matrix and timeseries data
try <- simu_var(method = 'sparse', nobs = T, k=p, sp_pattern= 'random', arlags=seq(1,q.t,1),
                brk=brk, sp_density = sp_density , seed = 1 )
#display the true AR coefficients matrice
print(plot.matrix(do.call("cbind",try$model_param), m ))
data <- try$series
data <- as.matrix(data)

```

## Block Segmentation Scheme (BSS) Implementation 

After data generataion, we now perform the block segmentation scheme (BSS) algorithm to detect the structural breaks in large scale high-dimensional non-stationary VAR models. The BSS algorithm mainly contains three steps:

* First step: initial break points selection by block fused lasso.
* Second step: local screening by minimizing a localized information criterion (LIC) to eliminate candidate break points that are located far from any true break points. 
* Third step: exhaustive search by comparing SSE for each candidate point in a given cluster and keeping one break point in each cluster. 

The proposed BSS algorithm in the manuscript is implemented by the function $\texttt{bss()}$. The only input argument is the time series data, with each column representing the time series component.  
```{r}
#run the bss method
temp <- bss(data)

#display the estimated break points
print("Estimated break points:")
print(temp$final.selected.points)
#display the true break points
print("True break points:")
print(brk[-length(brk)])
```

## Optional Arguments and Default Values

In addition to the argument $\texttt{data}$  which was described above, there are other optional arguments for the function $\texttt{bss()}$ which are listed below: 

* $\texttt{lambda.1.cv}$: the tuning parmaeter $\lambda_1$ for fused lasso. By default, $\texttt{lambda.1.cv}$ is a vector with decreasing values constructed based on the time series data, the number of time series components $\texttt{p}$, and the block size $\texttt{block.size}$.
* $\texttt{lambda.2.cv}$: the tuning parmaeter $\lambda_2$ for fused lasso. By default, $\texttt{lambda.2.cv}$ is a vector with decreasing values constructed based on the  number of time points $\texttt{T}$ and the number of time series components $\texttt{p}$. 
* $\texttt{q}$: the AR order. By default, $\texttt{q}$ = 1.
* $\texttt{max.iteration}$: the max number of iteration for the fused lasso. By default, $\texttt{max.iteration}$ = 50. 
* $\texttt{tol}$: tolerance for the fused lasso. By default, $\texttt{tol}$ = $10^{-2}$. 
* $\texttt{block.size}$: the block size $b_n$. By default, $b_n= \lfloor \sqrt{T} \rfloor$.
* $\texttt{blocks}$: the blocks (sequence). By default, $\texttt{blocks = seq(0,T,block.size)}$. One could also use $\texttt{blocks}$ to handle varying-size blocks. For example, $\texttt{blocks= c(seq(0,5000,100), seq(5200,10000,200))}$. 
* $\texttt{an.grid}$: a list of values for $a_n$. By default it is NULL.


